---
title: "Package Overview"
author: "Jonathan Kadish"
date: "3/22/2021"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Spatial Object packages
The basics are sf and sp
# sp
sp implements our basic spatial data types using spatial "classes"
```{r}
library(pacman)
p_load(sp)
getClass("Spatial")
```
# Loading data into an sp object
We can load any data with coordinates into the sp package. These can be points, lines, of polgons.
```{r}
coords <- rbind(c(1.5, 2), c(2.5, 2), c(0.5, 0.5), c(1, 0.25), c(1.5, 
    0), c(2, 0), c(2.5, 0), c(3, 0.25), c(3.5, 0.5))
pts <- SpatialPoints(coords)
plot(pts)
```
Note it is easy to use the plot function with a spatial classes, but there is also a ggspatial package to make them graphable with ggplot. However, I could not get this to work.
```{r, eval = FALSE}
p_load(ggspatial, ggplot2)
ggplot() +
  geom_spatial_point(data = pts)
```
# CRS
Remember, to analyze spatial data we need to specify the coordinate reference system.
```{r}
is.projected(pts)
crs.geo <- CRS("+init=EPSG:32633")  # UTM 33N
proj4string(pts) <- crs.geo  # define projection system of our data
is.projected(pts)
```
# Add data
We add attributes to a spatial class to make it a spatial points data frame
```{r}
df <- data.frame(attr1 = c("a", "b", "z", "d", "e", "q", "w", "r", "z"), attr2 = c(101:109))
df
spdf <- SpatialPointsDataFrame(pts, df)
summary(spdf)
```
# sf
sf (simple feature) is newer than sp and is growing in popularity, in part because it integrates nicely with the tidyverse suite. It's easy to transfer data from one to the other
```{r}
p_load(sf)
ptssf <- st_as_sf(pts)
class(ptssf)
st_geometry(ptssf)
#can still use the plot function
plot(ptssf)
#can also use ggplot
ggplot(ptssf) +
  geom_sf()
```
Why not always use sf? Some packages work with sf, some with sp. But we can convert from sf to sp object easily as well. Many packages now work with both.My recommendation: use sf except when you have to use sp.
```{r}
#convert sf back to sp
pts2 <- as(ptssf, "Spatial")
```
# Reading in shapefile
If you download a shapefile and want to read it in directly, use the raster package (shapefile) [relies on rgdal package], the readOGR (rgdal package) [more flexible], or st_read (sf package) [to read straight to a simple feature object].
# Spatial autocorrelation
One package that uses the sp data structure is the one we used for neighbors and spatial autocorrelation, spdep.
```{r}
us_states <- us_states
class(us_states)
p_load(spdep)
# spdep takes either sp or sf objects
nn <- poly2nb(us_states)
nn.list <- nb2listw(nn)
#one benefit of sp: plotting a neighbor object is very easy
us_sp <- as(us_states, "Spatial")
plot(us_sp)
plot(nn, coordinates(us_sp), add = TRUE)
```
For spatial autocorrelation, we used moran.test. We use the list version of our neighbor object
```{r}
# analytical version
moran.test(us_states$total_pop_10, listw = nn.list)
# monte carlo version
mc <- moran.mc(us_states$total_pop_10, listw = nn.list, nsim = 99)
mc
plot(mc)
```
# interpolation
We used a few different packages for interpolation. Nearest neighbor interpolatuion using the dismo package (voronoi command), gstat package (inverse distance weighting).
```{r}
p_load(dismo)
vsp <- voronoi(pts)
plot(vsp)
#note, this package only works with sp object, or a two column matrix of coordinates
vsp <- voronoi(ptssf)
```
```{r}
p_load(gstat)
#create gstat object
gs <- gstat(formula = attr2 ~ 1, data = spdf)
#to interpolate to a field, we create a raster then use the raster package's interpolate function.
r <- raster(spdf, res = .1)
ridw <- interpolate(r, gs)
plot(ridw)
```
# Point processes
Raster package was also useful for quadrats
```{r}
quad <- rasterize(coordinates(pts),fun = 'count', r)
plot(quad)
```
For point processes, we use the spatstat package. This package only works with sp objects
```{r}
library(spatstat)
#create a window
coordinates(pts)
#will automatically create a spatstat object from an sf object using as.ppp
pts.stat <- as.ppp(ptssf)
#from an sp object, you need to create the owin
#can create it manually or prom a polygon object as we did in class
pts.owin <- owin(c(0,4), c(0,4))
coords <- coordinates(pts)
pts.stat.sp <- ppp(x = coords[,1], y = coords[,2],window =  pts.owin)
```
With spatstat, it was easy to compute Kernel Density, F, G, K functions. Do that here:
```{r}
```
Collapse





Send a message to in-class















