---
title: "Practice with R"
author:
date: 
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.align = "center")
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)
```

---
# Practice with R
# Case Study: Sampling

We can estimate the mean $\mu$ of a population of size $N$ in two ways:

--

1. Take a simple random sample of size $n$ and calculate $\bar{x}$.

--

2. Take a stratified sample of $n_1$ from the first strata and $n_2$ from the 
second strata, proportionally, then compute $\bar{x}$.

--

Which estimate (r.v.) will have "better" properties?

---
# Set up population

```{r}
set.seed(300)
```

--

```{r}
# generate population
N_1 <- 30000
N_2 <- 40000
N_3 <- 50000
g1 <- rnorm(N_1, mean = 1, sd = 1)
g2 <- rnorm(N_2, mean = 2, sd = 1)
g3 <- rnorm(N_3, mean = 3, sd = 1)
pop <- c(g1, g2, g3)
```

--

```{r}
# lay out parameters
N <- N_1 + N_2 + N_3
l_1 <- N_1/N
l_2 <- N_2/N
l_3 <- N_3/N
mu_1 <- mean(g1)
sigsq_1 <- var(g1) * (N_1 - 1) / N_1
mu_2 <- mean(g2)
sigsq_2 <- var(g2) * (N_2 - 1) / N_2
mu_3 <- mean(g3)
sigsq_3 <- var(g3) * (N_3 - 1) / N_3
```


---
# 1. One simple random sample 

--

```{r fig.height=3, fig.align="center"}
n <- 100
samp <- sample(pop, size = n)
df <- data.frame(x = samp)
library(tidyverse)
ggplot(df, aes(x = x)) +
  geom_dotplot(alpha = .5) +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```


---
# 1. One simple random sample, cont 

```{r fig.height = 3}
# Density plot
ggplot(df, aes(x = x)) +
  geom_density(fill = "steelblue", alpha = .5) +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```


---
# 1. Simulation: SRS

--

```{r}
it <- 5000
xbar_srs <- rep(NA, it)
for (i in 1:it) {
  samp <- sample(pop, size = n)
  xbar_srs[i] <- mean(samp)
}
```

--

```{r}
head(xbar_srs)
```


---
# 2. Two stratified sample (With three strata)

--

```{r}
n_1 <- n * l_1
n_2 <- n * l_2
n_3 <- n * l_3
samp_1 <- sample(g1, size = n_1)
samp_2 <- sample(g2, size = n_2)
samp_3 <- sample(g3, size = n_3)
df <- data.frame(x = c(samp_1, samp_2, samp_3),
                 stratum = rep(c("1", "2", "3"), 
                               times = c(n_1, n_2, n_3)))
```

--

```{r, fig.height=3, echo = FALSE}
ggplot(df, aes(x = x, fill = stratum)) +
  geom_dotplot(alpha = .5) +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```


---
# 2. Simulation: Stratified sampling

--

```{r}
xbar_strat <- rep(NA, it)
for (i in 1:it) {
  samp_1 <- sample(g1, size = n_1)
  samp_2 <- sample(g2, size = n_2)
  samp_3 <- sample(g3, size = n_3)
  xbar_strat[i] <- mean(c(samp_1, samp_2, samp_3 ))
}
```

```{r}
head(xbar_strat)
```


---
# Simulations compared 

```{r, fig.height=3}
df <- data.frame(xbar = c(xbar_srs, xbar_strat),
                 method = rep(c("srs", "strat"), times = c(it, it)))
ggplot(df, aes(x = xbar, fill = method)) +
  geom_density(alpha = .5) +
  theme_bw() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
# Conclusion: From the comparision density plot, the plot for the stratified samples has a distribution with less variance, suggesting that the stratified sampling gives rvs with "better" properties than the random sampling. 
